---
title: "Project 1 Checkpoint 3 Report"
format: html
execute: 
  warning: false
  message: false
embed-resources: true
theme: darkly
---

# Source 1:

We are utilizing REST count

Importing data and libraries

```{r}
library(httr)
library(tidyjson)
library(jsonlite)
library(tidyverse)
library(glue)
library(countrycode)

tuesdata <- tidytuesdayR::tt_load('2021-01-26')
tuesdata <- tidytuesdayR::tt_load(2021, week = 5)
plastics <- tuesdata$plastics
```

Grabbing regions from the API and mapping them to preset labels

```{r}
#| output: false
get_region <- function(region) {
  url <- glue(
    "https://restcountries.com/v3.1/region/{region}?fields=name,population,region"
    )

  res <- GET(url)
  stop_for_status(res)

  txt <- content(res, as = "text", encoding = "UTF-8")
  df <- fromJSON(txt, flatten = TRUE)

  as_tibble(df) %>%
    transmute(
      country_api = coalesce(name.common, name.official),
      region = region,
      population = population
    )
}

region_list <- c(
  "africa",
  "americas", 
  "asia", 
  "europe", 
  "oceania"
  )

country_region_lookup <- map_dfr(region_list, get_region) %>%
  distinct() %>%
  filter(!is.na(country_api))

country_region_lookup
```

Cleaning up data

```{r}
#| output: false
plastic_data_clean <- plastics %>%
  mutate(
    country_clean = str_squish(country),
    country_clean = recode(
      country_clean,
      "USA" = "United States",
      "US" = "United States",
      "UK" = "United Kingdom",
      "Czech Republic" = "Czechia",
      "Republic of the Congo" = "Congo",
      "Democratic Republic of the Congo" = "DR Congo",
      .default = country_clean
    )
  )

lookup_clean <- country_region_lookup %>%
  mutate(
    country_clean = str_squish(country_api)
  )

plastic_with_region <- plastic_data_clean %>%
  left_join(
    lookup_clean %>% select(country_clean, region, population),
    by = "country_clean"
  )

unmatched_countries <- plastic_with_region %>%
  filter(is.na(region)) %>%
  distinct(country, country_clean)
```

Creating the table

```{r}
plastic_by_region <- plastic_with_region %>%
  filter(!is.na(region)) %>%
  group_by(region) %>%
  summarise(
    total_plastic = sum(grand_total, na.rm = TRUE),
    countries_included = n_distinct(country),
    total_population = sum(population, na.rm = TRUE),
    plastic_per_capita = total_plastic / total_population,
    .groups = "drop"
  ) %>%
  arrange(
    desc(
      total_plastic
      )
    )

print(plastic_by_region)
```

# Source #2: 

We are choosing to utilize API-Ninjas API to gather country GDP and determine whether there is a pattern between a countries GDP and their total number of volunteers.

```{r}
#| output: false
my_key = "k6XuPbvIYFC3YPIaD875hrv3Tr4WfWhFXUsyhuFQ"
get_GDP <- function (country, key = my_key) {
  
  url <- glue(
    "https://api.api-ninjas.com/v1/gdp?country={country}&year=;X-Api-Key={my_key}"
    )
  my_data <- safely(fromJSON)(url)
  if (is.null(my_data$result)) {
    return(NA)
  } else {
    return(my_data$result)
  }
}
countries <-unique(plastics$country)
results <- map(
  .x = countries,
  .f = get_GDP
  )
```

Finding missing countries in the API and using manual mapping of 2-letter country codes.

```{r}
#| output: false
bad_countries <- countries[map_lgl(results, ~ length(.x) == 0 || (length(.x) == 1 && is.na(.x)))]

bad_country_code <- c(
  "BF",  
  "CI", 
  "HK",  
  "ZA",
  "LK",
  "TW",
  "AE",
  "GB",
  "US",
  "SV",
  "KR",
  "GB"
)

results_2 <- map(
  .x = bad_country_code, 
  .f = get_GDP
  )

mod_df <- plastics[!(plastics$country %in% bad_countries), ]
unique(plastics$country)

results <- map(
  .x = unique(mod_df$country), 
  .f = get_GDP
  )

total_results <- bind_rows(
  results, 
  results_2
  )

tr_2019 <- total_results[total_results$year %in% c(2019), ]
```

Combining good and bad countries and plotting the data.

```{r}

plastics_2019_sum <- plastics |>
  filter(year == 2019) |>
  group_by(country) |>
  summarize(total_volunteers = sum(volunteers, na.rm = TRUE)) |>
  mutate(iso = countrycode(country, "country.name", "iso3c"))

combined_2019 <- inner_join(
  plastics_2019_sum,
  tr_2019,
  by = c("iso" = "country")
)

combined_2019 <- unique(combined_2019)
write.csv(combined_2019, "plotting-data.csv")
options(scipen = 999) #to turn off scientific notation on graph
plot(combined_2019$total_volunteers ~ combined_2019$gdp_per_capita_nominal, 
     col = "blue", 
     pch = 16, 
     xlab = "GDP per capita (nominal, 2019)", 
     ylab = "Total volunteers",
     main = "Volunteers vs GDP per Capita (nominal, 2019)", log = "x"
     )
```